EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-08-21 draft.tex
6.3 revised; please @mariosky check 1 comment

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. In this section, we give specific details
about how we deployed EvoSwarm in a cloud-setting for running the experiments
within this paper. Running the cloud-native application in the cloud has the
benefit of a multi-instance deployment, where we can run containers in several
instances executing at the same time, not only in a single machine.

When we deploy a cloud-native application to a cloud provider, we have two
options to choose:

@@

  \item[A Kubernetes cluster] provides auto-scaling and management of
  containerized applications, but it needs additional configuration for
  provisioning. All major cloud-providers have Kubernetes services. This option
  was used by S. Zhao et al. \cite{zhao2019cloud} to run a multi-island
  algorithm, and Dziurzanski et al. \cite{dziurzanski2020scalable} later refined
  it.

  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.
    Salza et al. \cite{salza2019speed} tested this option by using CoreOs cluster.

@@

% Remember to mention cost at the end - JJ 
% Ping - JJ

We can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
AWS Fargate service. Fargate is a serverless compute engine for containers that
provides automatic provision and management of AWS ECS resources. With this
service, calculating the cost of execution is simplified because users only
specify the computing units and memory for the application, instead of every
component. However, AWS Fargate is not suited for an experimental setting because
many of the details are out of user control. For this reason, we chose the
alternative of using a cluster of EC2 containers. The drawback is that we need to manually
provision the  EC2 containers, networking, and group permissions for each
experimental setup, for instance, when going from 2 to 8 workers, we need to increase the
number of EC2 instances to meet the required amount of CPU units.

Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
parallel. Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can also specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt cpu\_shares} indicating the CPU units
assigned, in this case 2048; and the network configuration, we have two private 
subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
need to be supported by the currently available cluster. For instance, we need to have all 
the CPU units required by all tasks, if they are not available, the task initiation will fail.

@@

Each container is defined as a {\texttt service} (task in AWS terms) , with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
variables used by code running inside containers.

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws_ec2}
  \caption{AWS EC2 configuration for the deployment of EvoSwarm containers.
   }
  \label{fig:aws-configuration}
\end{figure}

Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
of experiments. We selected three types of EC2 instances according to our needs, the  
{\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
and we do not need to scale them as the amount of parallel workers are increased. They run in a
single {\texttt C5  High-CPU large} instance. With the available 2048 CPU shares divided as follows:
{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). With this assignment 
the {\texttt redis} container gets a dedicated vCPU, needed because it is the most demanding 
task of the three. We assigned 2048 units to each {\texttt worker} container, and using a more 
capable {\texttt C5 High-CPU XLarge} instance,  with 
four vCPUs and 4096 CPU shares, with this configuration each worker gets a pair of vCPUs.  
We scaled the number if instances as needed, using one instance for every two {\texttt worker} containers.

Because all these instances run in a private subnet, we also
run a Bastion instance in a public subnet, from where we send the commands to execute the experiments 
and later collected the results. We connected to this instance through a {\texttt ssh} session from 
our personal workstation. This is a {\texttt R5AD Large} instance, selected with 75GiB of storage, needed for
keeping the logs, and more memory (16GiB) for processing the logs in memory. The details these EC2 instances 
and their cost at the time of writing are shown in Table \ref{tab:ec2}.

>>>>

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. In this section, we give specific details
about how we deployed EvoSwarm in a cloud-setting for running the experiments
within this paper. Running the cloud-native application in the cloud has the obvious
benefit of a multi-instance deployment, where we can run containers in several
instances executing at the same time, not only in a single machine.

When we deploy a cloud-native application to a cloud provider, there
are, among others, two options to choose from:

@@

  \item[A Kubernetes cluster] provides auto-scaling and management of
  containerized applications, but it needs additional configuration for
  provisioning and its cost might be higher. All major cloud-providers have Kubernetes services. This option
  was used by S. Zhao et al. \cite{zhao2019cloud} to run a multi-island
  algorithm, and Dziurzanski et al. \cite{dziurzanski2020scalable} later refined
  it.

  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift
    (which is also open source), and Marantis Docker Enterprise.
    Salza et al. \cite{salza2019speed} tested this option by using CoreOs cluster.

@@

We can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
AWS Fargate service. Fargate is a serverless compute engine for containers that
provides automatic provision and management of AWS ECS resources. With this
service, calculating the cost of execution is simplified because users only
specify the computing units and memory for the application, instead of every
component. However, AWS Fargate is not suited for an experimental setting because
many of the details are out of user control. For this reason, we chose the
alternative of using a cluster of EC2 containers. The drawback is that we need to manually
provision the  EC2 containers, networking, and group permissions for each
experimental setup; for instance, when going from 2 to 8 workers, we need to increase the
number of EC2 instances to meet the required amount of CPU units.

Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
parallel. Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can also specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt cpu\_shares} indicating the CPU units
assigned, in this case $2048$, and the network configuration; in this case we have two private 
subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
need to be supported by the currently available cluster. For instance, we need to have all 
the CPU units required by all tasks, if they are not available, the task initiation will fail.

@@

Each container is defined as a {\texttt service} ({\em task} in AWS terms), with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
variables used by code running inside containers.

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws_ec2}
  \caption{AWS EC2 configuration for the deployment of EvoSwarm containers.
  }
% Add authorship if necessary - JJ
  \label{fig:aws-configuration}
\end{figure}

Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
of experiments. We selected three types of EC2 instances according to our needs; the  
{\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
and we do not need to scale them as the amount of parallel workers are increased. They run in a
single {\texttt C5  High-CPU large} instance. With the available 2048 CPU shares divided as follows:
{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). With this assignment 
the {\texttt redis} container gets a dedicated vCPU, needed because it is the most demanding 
task of the three. We assigned 2048 units to each {\texttt worker} container, and using a more 
capable {\texttt C5 High-CPU XLarge} instance,  with 
four vCPUs and 4096 CPU shares, with this configuration each worker gets a pair of vCPUs.  
We scaled the number if instances as needed, using one instance for every two {\texttt worker} containers.

Because all these instances run in a private subnet, we also
run a Bastion instance in a public subnet, from where we send the
% what is a Bastion instance? - JJ
commands to execute the experiments
and later collected the results. We connected to this instance through a {\texttt ssh} session from 
our personal workstation. This is a {\texttt R5AD Large} instance, selected with 75GiB of storage, needed for
keeping the logs, and more memory (16GiB) for processing the logs in memory. The details these EC2 instances 
and their cost at the time of writing are shown in Table \ref{tab:ec2}.
